---
title: "Solar Project"
author: "Rose Dennis"
date: "3/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, results='hide'}
library(lubridate)
library(ggplot2)
library(tidyverse)
library(car)
library(dplyr)
library(lattice)
library(MASS)
library("ISLR", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
library("glmnet", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
library("randomForest", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
library("pROC", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
library("pls", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
library("corrplot", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
```

GOALS:
-       Question: Solar power generation (gen) based other variables (just 
consumption of power covariates)
-       Check for linear relationship between variables
-       Finding out which variables are significant to the model by performing 
various regression techniques (pairs plot)
-       Check assumptions


```{r}
setwd("/Users/rosedennis/Desktop/Applied Math Grad School/ProjectSeminar")
```

GETTING MULTIPLOT FUNCTION TO WORK
```{r, results='hide'}
multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)

  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                 ncol = cols, nrow = ceiling(numPlots/cols))
}

if (numPlots == 1) {
print(plots[[1]])

} else {
grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

for (i in 1:numPlots) {
  matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

  print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                  layout.pos.col = matchidx$col))
 }
}
 }
```


```{r}
#multiplot(gen.hist, hum.hist, appTEMP.hist, cc.hist, dewPOINT.hist, temp.hist, cols=2)
#multiplot(vis.hist, press.hist, windBearing.hist, precipProb.hist, windSpeed.hist, pricipIntensity.hist, cols=2)
```

for solar generation want regression for afternoon for each season (4 diff models)---afternoon hour 13-17

NEW DATASETS (JUST PEAK HOURS)

```{r}
peak2014<- dplyr::select(homeC2014DAY,
                  temp:gen,
                  season
)
peak2014winter<-filter(peak2014,season=="winter")
peak2014summer<-filter(peak2014,season=="summer")
peak2014fall<-filter(peak2014,season=="fall")
peak2014spring<-filter(peak2014,season=="spring")
```

WINTER

Descriptives
```{r}
#Response Histogram
gen.hist <- ggplot(data=peak2014winter, aes(x=gen)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Solar Generation") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
gen.hist
#doesn't look great, tried log transformation
log.gen.hist<- ggplot(data=peak2014winter, aes(x=log(gen))) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Log(Winter Solar Generation)") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#still not great but oh well
#Predictor Histograms (just continuous variables)
#HUMIDITY
hum.hist <- ggplot(data=peak2014winter, aes(x=hum)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Humidity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#tried log transform, looks worse
#AppTEMP
appTEMP.hist <- ggplot(data=peak2014winter, aes(x=appTemp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter appTEMP") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#looks normal
#CC
cc.hist <- ggplot(data=peak2014winter, aes(x=cc)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter CC") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#logtransorm...idk if this looks better
log.cc.hist<- ggplot(data=peak2014winter, aes(x=log(cc))) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Log(CC)") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#dewPOINT
dewPOINT.hist <- ggplot(data=peak2014winter, aes(x=dewPoint)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Dew Point") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#TEMP
temp.hist <- ggplot(data=peak2014winter, aes(x=temp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Temperature") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#VIS
vis.hist <- ggplot(data=peak2014winter, aes(x=vis)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Visibility") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#looks bad but log transform doesn't help
#PRESS
press.hist <- ggplot(data=peak2014winter, aes(x=press)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Pressure") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windBEARING
windBearing.hist <- ggplot(data=peak2014winter, aes(x=windBearing)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Wind Bearing") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#looks bimodal, not sure how to fix that
#precipPROB
precipProb.hist <- ggplot(data=peak2014winter, aes(x=precipProb)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Precipitation Probability") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windSPEED
windSpeed.hist <- ggplot(data=peak2014winter, aes(x=windSpeed)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Wind Speed") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#skewed slightly right
#pricipINTENSITY
pricipIntensity.hist <- ggplot(data=peak2014winter, aes(x=pricipIntensity)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Pricipitation Intensity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#tried log transform, idk what looks better
#USE
use.hist <- ggplot(data=peak2014winter, aes(x=use)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Winter Use") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#skewed right
multiplot(gen.hist, hum.hist, appTEMP.hist, cc.hist, dewPOINT.hist, temp.hist, vis.hist, press.hist, windBearing.hist, precipProb.hist, windSpeed.hist, pricipIntensity.hist, cols=2)
```

MODEL-NO TRANSFORMS

```{r}
peak2014winter_2<- dplyr::select(peak2014winter,temp:gen)
peak2014winter_3<-na.omit(peak2014winter_2) #only uses complete sets of data. This decreased the nunber of observations by a lot but it was the only way I could make the code work. went from 810 obs to 586
w1<-lm(gen~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb, data=peak2014winter_3)
summary(w1) #R^2 is kind of low = .3169
qqnorm(w1$residuals) #qq plot looks fine
anova(w1)
#says temp,icon,hum,vis,press, and windBearing are important... R^2=.3203

#Using AIC to choose variables
both1 = step(w1, direction = "both", trace=FALSE ) 
summary(both1) #temp, iconclear-night, iconcloudy, press, windBearing, dewPoint, precipProb ...R^2=.3238
backwards1 = step(w1, direction = "backward", trace=FALSE ) 
summary(backwards1) #temp, iconclear-night, iconcloudy, press, windBearing, dewPoint, precipProb ...R^2=.3238
forward1 = step(w1, direction = "forward", trace=FALSE ) 
summary(forward1) #iconclear-night, press, windBearing, ...R^2=.3169
#going to keep temp, press, cc, windBearing, dewPoint and use

#backwards selection best

w_final<-lm(gen~temp+icon+press+windBearing+dewPoint+precipProb, data=peak2014winter_3)
summary(w_final) #R^2=.3238

#performed a pairs plot to look at correlation. Had to take out icon because it's discrete. Also calculated Variance Inflation Factors for multicollinearity.
cor_df<-dplyr::select(peak2014winter_3, temp, hum, vis, appTemp, press, windSpeed, cc, windBearing, pricipIntensity, dewPoint, precipProb, gen)
M<-cor(cor_df)
corrplot(M, method= "number")
vif(w_final) #idt any of these have a really high VIF so we're okay
qqnorm(w_final$residuals)
#CHECKING ASSUMPTIONS
#the residuals vs fitted plot looks really bad...they clearly have some kind of trend
errors1<-ggplot(fortify(w_final), aes(x=.fitted, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Fitted Values")+ylab("Residuals") +
      ggtitle("Residual vs Fitted Plot")
#check for non-constant variance
variance1<-ggplot(data=w_final, aes(x=.resid)) + 
      geom_histogram(binwidth=0.5, fill="blue") +
      xlab("Residuals") +
      ggtitle("Distribution of Residuals")
```

Test MSE

```{r}
set.seed(11)
train<-sample(1:dim(peak2014winter_3)[1],dim(peak2014winter_3)[1]/2) #splitting data into half train.
test<- -train
Winter.train<- peak2014winter_3[train,] 
Winter.test<- peak2014winter_3[test,] 

#linear model
w_a<-lm(gen~temp+icon+press+windBearing+dewPoint+precipProb, data=Winter.train)
pred.fit<-predict(w_a, Winter.test)
MSE_linear<-mean((pred.fit-Winter.test$gen)^2) #test MSE is 0.0162
MSE_linear

#ridge
x<-model.matrix(gen~.-use,peak2014winter_3 )[,-1] #the normal design matrix has a column of ones so we get rid of that,
y<-peak2014winter_3$gen 
y.test<-y[test] 
cv.out<-cv.glmnet (x[train ,],y[train],alpha =0) #alpha=0 is ridge 
plot(cv.out)
bestlam<-cv.out$lambda.min
bestlam #this seems low...lambda=0.0295
ridge.mod<-glmnet(x[train,],y[train],alpha =0, lambda =bestlam , 
thresh =1e-12)
ridge.pred=predict(ridge.mod ,s=bestlam, newx=x[test ,])
MSE_ridge<-mean((ridge.pred -y.test)^2) #test MSE is 0.0153
MSE_ridge

#lasso
cv.out2<-cv.glmnet (x[train ,],y[train],alpha =1) #alpha=1 for lasso #same error as above
plot(cv.out2)
bestlam2<-cv.out2$lambda.min
bestlam2 #lambda=0.0045
lasso.mod<-glmnet(x[train,],y[train],alpha =1, lambda =bestlam2 , 
thresh =1e-12)
lasso.pred=predict(lasso.mod ,s=bestlam2, newx=x[test ,])
MSE_lasso<-mean((lasso.pred -y.test)^2) #test MSE=0.0150
MSE_lasso
out=glmnet(x,y,alpha =1, lambda =bestlam2)
lasso.coef=predict(out ,type ="coefficients",s=bestlam )[1:18 ,]
lasso.coef #does variable selection, says there are 5 non-zero coeffs- iconpartly-cloudy-night, hum, temp, vis, iconclear-night

#PCR
pcr.fit=pcr(gen~.-use, data=peak2014winter_3 ,subset =train ,scale =FALSE , #had to change scale to false or else it didn't work
validation ="CV")
validationplot(pcr.fit,val.type="MSEP")

#Looking at the plot, it seems that the lowest cross validation error is around M=8. So we will use this in finding the test MSE.

pcr.pred=predict (pcr.fit ,x[test ,], ncomp =8)
MSE_pcr<-mean((pcr.pred -y.test)^2) #test MSE=0.0152
MSE_pcr

#PLS
pls.fit=plsr(gen~.-use, data=peak2014winter_3 ,subset =train ,scale=FALSE ,
validation ="CV")
validationplot(pls.fit ,val.type="MSEP") #looks like the lowest point is around 8
pls.pred=predict (pls.fit ,x[test ,], ncomp =8)
MSE_pls<-mean((pls.pred -y.test)^2) #test MSE=0.0152
MSE_pls

#comparison
test.avg<- mean(Winter.test$gen)
lmodel2<-1-mean((pred.fit-Winter.test$gen)^2)/mean((test.avg-Winter.test$gen)^2)
r2<-1-mean((ridge.pred-Winter.test$gen)^2)/mean((test.avg-Winter.test$gen)^2)
l2<-1-mean((lasso.pred-Winter.test$gen)^2)/mean((test.avg-Winter.test$gen)^2)
pcr2<-1-mean((pcr.pred-Winter.test$gen)^2)/mean((test.avg-Winter.test$gen)^2)
pls2<-1-mean((pls.pred-Winter.test$gen)^2)/mean((test.avg-Winter.test$gen)^2)
c(lmodel2,r2,l2,pcr2,pls2) #PLS performs the best

#decision tree
m_tree<-tree(gen~.-use, data=Winter.train)
summary(m_tree) #used temp, windSpeed, press, hum, icon, appTemp, vis, windBearing, and dewPoint 
m_tree #15 terminal nodes, deviance=sum of squared errors= 0.0086
plot(m_tree)
text(m_tree,pretty=0)

cv.tree =cv.tree(m_tree)
plot(cv.tree$size ,cv.tree$dev ,type='b') #I think this shows that the tree size that best minimizes the cv error is a tree of size 3

prune.tree =prune.tree(m_tree ,best =3) #3 terminal nodes, uses temp and windSpeed deviance=0.01408
plot(prune.tree )
text(prune.tree,pretty =0) 

yhat=predict (m_tree,newdata =Winter.test)
plot(yhat ,Winter.test$gen)
abline (0,1)
MSE_tree<-mean((yhat -Winter.test$gen)^2) #test MSE= 0.0194
MSE_tree
sqrt(mean((yhat -Winter.test$gen)^2)) #solar predictions within around 0.1394 of true median generation value for the winter

yhat.prune=predict (prune.tree,newdata =Winter.test)
plot(yhat.prune ,Winter.test$gen)
abline (0,1)
MSE_prune<-mean((yhat.prune -Winter.test$gen)^2) #test MSE= 0.0174
MSE_prune
sqrt(mean((yhat.prune -Winter.test$gen)^2)) #solar predictions within around 0.1320 of true median generation value for the winter....this is better

#random forest  
rf =randomForest(gen~.-use,data=peak2014winter_3 ,subset =train ,
mtry=5, importance =TRUE) #used mtry=5 because it's usually p/3 for regression
yhat.rf = predict (rf,newdata =Winter.test)
importance(rf)
varImpPlot (rf)
MSE_rf<-mean(( yhat.rf -Winter.test$gen)^2) #test MSE=0.0081, performs the best
MSE_rf

```

SUMMER

Descriptives
```{r}
gen.hist3 <- ggplot(data=peak2014summer, aes(x=gen)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Solar Generation") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#Predictor Histograms (just continuous variables)
#HUMIDITY
hum.hist3 <- ggplot(data=peak2014summer, aes(x=hum)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Humidity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#AppTEMP
appTEMP.hist3 <- ggplot(data=peak2014summer, aes(x=appTemp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer appTEMP") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#CC
cc.hist3 <- ggplot(data=peak2014summer, aes(x=cc)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer CC") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#dewPOINT
dewPOINT.hist3 <- ggplot(data=peak2014summer, aes(x=dewPoint)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Dew Point") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#TEMP
temp.hist3 <- ggplot(data=peak2014summer, aes(x=temp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Temperature") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#VIS
vis.hist3 <- ggplot(data=peak2014summer, aes(x=vis)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer  Visibility") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#PRESS
press.hist3 <- ggplot(data=peak2014summer, aes(x=press)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Pressure") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windBEARING
windBearing.hist3 <- ggplot(data=peak2014summer, aes(x=windBearing)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Wind Bearing") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#precipPROB
precipProb.hist3 <- ggplot(data=peak2014summer, aes(x=precipProb)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Precipitation Probability") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windSpeed
windSpeed.hist3 <- ggplot(data=peak2014summer, aes(x=windSpeed)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Wind Speed") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#pricipINTENSITY
pricipIntensity.hist3 <- ggplot(data=peak2014summer, aes(x=pricipIntensity)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Pricipitation Intensity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#USE
use.hist3 <- ggplot(data=peak2014summer, aes(x=use)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Summer Use") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
multiplot(gen.hist3, hum.hist3, appTEMP.hist3, cc.hist3, dewPOINT.hist3, temp.hist3, vis.hist3, press.hist3, windBearing.hist3, precipProb.hist3, windSpeed.hist3, pricipIntensity.hist3, cols=2)
```

MODEL
```{r}
peak2014summer_2<- dplyr::select(peak2014summer,temp:gen)
peak2014summer_3<-na.omit(peak2014summer_2)
summer1<-lm(gen~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb, data=peak2014summer_3)
summary(summer1) 
qqnorm(summer1$residuals)
anova(summer1) #keeps temp, icon, sum, appTemp, cc, windBearing, dewPoint, precipProb... R^2=.2866
both3 = step(summer1, direction = "both", trace=FALSE ) 
summary(both3) #iconclear-night,iconpartly-cloudy-night,hum,sumLight Rain, sumMostly Cloudy, appTemp, cc, windBearing, pricipIntensity dewPoint, precipProb ...R^2=.2959
backwards3 = step(summer1, direction = "backward", trace=FALSE ) 
summary(backwards3) #iconclear-night,iconpartly-cloudy-night,hum,sumLight Rain, sumMostly Cloudy, appTemp, cc, windBearing, pricipIntensity, dewPoint, precipProb ...R^2=.2959
forward3 = step(summer1, direction = "forward", trace=FALSE ) 
summary(forward3) #iconclear-night,iconpartly-cloudy-night,hum,sumLight Rain, sumMostly Cloudy, cc, windBearing, pricipIntensity dewPoint, precipProb ...R^2=.2953

summer_final<-lm(gen~icon+hum+sum+appTemp+cc+ windBearing+pricipIntensity+ dewPoint+ precipProb, data=peak2014summer_3)
cor_df3<-dplyr::select(peak2014summer_3, temp, hum, vis, appTemp, press, windSpeed, cc, windBearing, pricipIntensity, dewPoint, precipProb, gen)
M3<-cor(cor_df3)
corrplot(M3, method= "number")
#vif(summer_final) #can't perform this b/c there seems to be perfectly correlated variables in the model. 
#run an alias to find these variables
alias( summer_final) #this shows that sumFoggy/iconfog, sumOvercast/iconcloudy, sumPartly Cloudy/iconpartly-cloudy-night, sumRain/iconrain, sumPartly Cloudy/sumMostly Cloudy, sumRain/sumDrizzle, sumRain/sumLight Rain are highly correlated. This makes sense just looking at the variable names.
qqnorm(summer_final$residuals)

#CHECKING ASSUMPTIONS
#the residuals vs fitted plot looks really bad...they clearly have some kind of trend
errors3<-ggplot(fortify(summer_final), aes(x=.fitted, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Fitted Values")+ylab("Residuals") +
      ggtitle("Residual vs Fitted Plot")
#check for non-constant variance
variance3<-ggplot(data=summer_final, aes(x=.resid)) + 
      geom_histogram(binwidth=0.5, fill="blue") +
      xlab("Residuals") +
      ggtitle("Distribution of Residuals")
```

Test MSE

```{r}
set.seed(11)
train3<-sample(1:dim(peak2014summer_3)[1],dim(peak2014summer_3)[1]/2) #splitting data into half train.
test3<- -train3
Summer.train<- peak2014summer_3[train3,] 
Summer.test<- peak2014summer_3[test3,] 
#linear model
summer_a<-lm(gen~icon+hum+sum+appTemp+cc+ windBearing+pricipIntensity+ dewPoint+ precipProb, data=Summer.train) #needed to take icon out for this
pred.fit3<-predict(summer_a, Summer.test)
MSE_linear3<-mean((pred.fit3-Summer.test$gen)^2) #test MSE is 0.0199
MSE_linear3
#ridge
x3<-model.matrix(gen~ .-use,peak2014summer_3 )[,-1] #the normal design matrix has a column of ones so we get rid of that,
y3<-peak2014summer_3$gen 
y.test3<-y3[test3] 
cv.out3<-cv.glmnet (x3[train3 ,],y3[train3],alpha =0) #alpha=0 is ridge 
plot(cv.out3)
bestlam3<-cv.out3$lambda.min
bestlam3 #this seems low...lambda=0.0050
ridge.mod3<-glmnet(x3[train3,],y3[train3],alpha =0, lambda =bestlam3 , 
thresh =1e-12)
ridge.pred3=predict(ridge.mod3 ,s=bestlam3, newx=x3[test3 ,])
MSE_ridge3<-mean((ridge.pred3 -y.test3)^2)#test MSE is 0.0186
MSE_ridge3
#lasso
cv.out3<-cv.glmnet (x3[train3 ,],y3[train3],alpha =1) #alpha=1 for lasso #same error as above
plot(cv.out3)
bestlam3<-cv.out3$lambda.min
bestlam3 #lambda=0.0025
lasso.mod3<-glmnet(x3[train3,],y3[train3],alpha =1, lambda =bestlam3 , 
thresh =1e-12)
lasso.pred3=predict(lasso.mod3 ,s=bestlam3, newx=x3[test3 ,])
MSE_lasso3<-mean((lasso.pred3 -y.test3)^2) #test MSE=0.0186
MSE_lasso3
out3=glmnet(x3,y3,alpha =1, lambda =bestlam3)
lasso.coef3=predict(out3 ,type ="coefficients",s=bestlam3 )[1:18 ,]
lasso.coef3 #iconpartly-cloudy-night, sumPartly Cloudy, sumMostly Cloudy, sumRain, iconclear-night, appTemp
#PCR
pcr.fit3=pcr(gen~.-use, data=peak2014summer_3 ,subset =train3 ,scale =FALSE , #had to change scale to false or else it didn't work
validation ="CV")
validationplot(pcr.fit3,val.type="MSEP")

#Looking at the plot, it seems that the lowest cross validation error is around M=10. So we will use this in finding the test MSE.

pcr.pred3=predict (pcr.fit3 ,x3[test3 ,], ncomp =10)
MSE_pcr3<-mean((pcr.pred3 -y.test3)^2) #test MSE=0.0214
MSE_pcr3
#pls
pls.fit3=plsr(gen~.-use, data=peak2014summer_3 ,subset =train3 ,scale=FALSE ,
validation ="CV")
validationplot(pls.fit3 ,val.type="MSEP") #looks like the lowest point is around 18 &...how do I get this more exact?
pls.pred3=predict (pls.fit3 ,x3[test3 ,], ncomp =10)
MSE_pls3<-mean((pls.pred3 -y.test3)^2) #test MSE= 0.0190
MSE_pls3

#decision tree
m_tree3<-tree(gen~.-use, data=Summer.train)
summary(m_tree3) #used windBearing, cc, press, dewPoint, pricipIntensity, and hum
m_tree3 #16 terminal nodes, deviance=sum of squared errors= 0.01389
plot(m_tree3)
text(m_tree3,pretty=0)

cv.tree3 =cv.tree(m_tree3)
plot(cv.tree3$size ,cv.tree3$dev ,type='b') #I think this shows that the tree size that best minimizes the cv error is a tree of size 10

prune.tree3 =prune.tree(m_tree3 ,best =10) #10 terminal nodes, uses windBearing, cc pricipIntensity, hum, deviance= 0.01568
plot(prune.tree3 )
text(prune.tree3,pretty =0) 

yhat3=predict (m_tree3,newdata =Summer.test)
plot(yhat3 ,Summer.test$gen)
abline (0,1)
MSE_tree3<-mean((yhat3 -Summer.test$gen)^2) #test MSE= 0.0236
MSE_tree3
sqrt(mean((yhat3 -Summer.test$gen)^2)) #solar predictions within around 0.1537 of true median generation value for the winter

yhat.prune3=predict (prune.tree3,newdata =Summer.test)
plot(yhat.prune3 ,Summer.test$gen)
abline (0,1)
MSE_prune3<-mean((yhat.prune3 -Summer.test$gen)^2) #test MSE= 0.0250
MSE_prune3
sqrt(mean((yhat.prune3 -Summer.test$gen)^2)) #solar predictions within around 0.1583 of true median generation value for the winter....this is better

#random forest  

rf3 =randomForest(gen~.-use,data=peak2014summer_3 ,subset =train3 ,
mtry=5, importance =TRUE)
yhat.rf3 = predict (rf3,newdata =Summer.test)
importance(rf3)
varImpPlot (rf3)
MSE_rf3<-mean(( yhat.rf3 -Summer.test$gen)^2) #test MSE=0.0144, performs the best
MSE_rf3

```

FALL

Descriptives

```{r}
#Response Histogram
gen.hist2 <- ggplot(data=peak2014fall, aes(x=gen)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Solar Generation") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
gen.hist2
#Predictor Histograms (just continuous variables)
#HUMIDITY
hum.hist2 <- ggplot(data=peak2014fall, aes(x=hum)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Humidity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#AppTEMP
appTEMP.hist2 <- ggplot(data=peak2014fall, aes(x=appTemp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall appTEMP") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#CC
cc.hist2 <- ggplot(data=peak2014fall, aes(x=cc)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall CC") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#dewPOINT
dewPOINT.hist2 <- ggplot(data=peak2014fall, aes(x=dewPoint)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Dew Point") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#TEMP
temp.hist2 <- ggplot(data=peak2014fall, aes(x=temp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Temperature") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#VIS
vis.hist2 <- ggplot(data=peak2014fall, aes(x=vis)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Visibility") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#PRESS
press.hist2 <- ggplot(data=peak2014fall, aes(x=press)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Pressure") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windBEARING
windBearing.hist2 <- ggplot(data=peak2014fall, aes(x=windBearing)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Wind Bearing") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#precipPROB
precipProb.hist2 <- ggplot(data=peak2014fall, aes(x=precipProb)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Precipitation Probability") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windSPEED
windSpeed.hist2 <- ggplot(data=peak2014fall, aes(x=windSpeed)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Wind Speed") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#pricipINTENSITY
pricipIntensity.hist2 <- ggplot(data=peak2014fall, aes(x=pricipIntensity)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Pricipitation Intensity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#USE
use.hist2 <- ggplot(data=peak2014fall, aes(x=use)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Fall Use") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
multiplot(gen.hist2, hum.hist2, appTEMP.hist2, cc.hist2, dewPOINT.hist2, temp.hist2, vis.hist2, press.hist2, windBearing.hist2, precipProb.hist2, windSpeed.hist2, pricipIntensity.hist2, cols=2)
```

MODEL- NO TRANSFORMS

```{r}
f1<-lm(gen~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb+use, data=peak2014fall)
summary(f1) 
qqnorm(f1$residuals)
anova(f1) #keep temp, icon, hum, vis, appTemp, cc, windBearing,dewPoint,precipProb, use ...R^2=.2929
both2 = step(f1, direction = "both", trace=FALSE ) 
summary(both2) #temp, iconclear-night,iconfog,iconpartly-cloudy-day,iconpartly-cloudy-night,hum,appTemp, cc, windBearing, dewPoint, precipProb, use ...R^2=.2932
backwards2 = step(f1, direction = "backward", trace=FALSE ) 
summary(backwards2) #temp, iconclear-night,iconfog,iconpartly-cloudy-day,iconpartly-cloudy-night,hum,appTemp, cc, windBearing, dewPoint, precipProb, use ...R^2=.2932
forward2 = step(f1, direction = "forward", trace=FALSE ) 
summary(forward2) #temp, iconclear-night,iconpartly-cloudy-day,iconpartly-cloudy-night,iconwind, sumBreezy and Mostly Cloudy, sumBreezy and Partly Cloudy,cc, windBearing, dewPoint, precipProb, use ... R^2=.2968

#FORWARD SELECTION WORKS BEST BUT IT HAS MANY VARIABLES KEPT IN THE MODEL. we include icon and sum
f_final<-lm(gen~temp+sum+cc+ windBearing+ dewPoint+ precipProb+ use, data=peak2014fall)
pairs(gen~temp+cc+windBearing+dewPoint+ precipProb+use, data=peak2014fall) #had to take out sum bc it's discrete
vif(f_final) #seems like sum and precipProb have high VIF
qqnorm(f_final$residuals)

#CHECKING ASSUMPTIONS
#the residuals vs fitted plot looks really bad...they clearly have some kind of trend
errors2<-ggplot(fortify(f_final), aes(x=.fitted, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Fitted Values")+ylab("Residuals") +
      ggtitle("Residual vs Fitted Plot")
#check for non-constant variance
variance2<-ggplot(data=f_final, aes(x=.resid)) + 
      geom_histogram(binwidth=0.5, fill="blue") +
      xlab("Residuals") +
      ggtitle("Distribution of Residuals")
```


SPRING

Descriptives
```{r}
gen.hist4 <- ggplot(data=peak2014spring, aes(x=gen)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Solar Generation") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#Predictor Histograms (just continuous variables)
#HUMIDITY
hum.hist4 <- ggplot(data=peak2014spring, aes(x=hum)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Humidity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#AppTEMP
appTEMP.hist4 <- ggplot(data=peak2014spring, aes(x=appTemp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring appTEMP") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#CC
cc.hist4 <- ggplot(data=peak2014spring, aes(x=cc)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring CC") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#dewPOINT
dewPOINT.hist4 <- ggplot(data=peak2014spring, aes(x=dewPoint)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Dew Point") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#TEMP
temp.hist4 <- ggplot(data=peak2014spring, aes(x=temp)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Temperature") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#VIS
vis.hist4 <- ggplot(data=peak2014spring, aes(x=vis)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Visibility") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#PRESS
press.hist4 <- ggplot(data=peak2014spring, aes(x=press)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Pressure") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windBEARING
windBearing.hist4 <- ggplot(data=peak2014spring, aes(x=windBearing)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Wind Bearing") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#precipPROB
precipProb.hist4 <- ggplot(data=peak2014spring, aes(x=precipProb)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Precipitation Probability") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#windSpeed
windSpeed.hist4 <- ggplot(data=peak2014spring, aes(x=windSpeed)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Wind Speed") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#pricipINTENSITY
pricipIntensity.hist4 <- ggplot(data=peak2014spring, aes(x=pricipIntensity)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Pricipitation Intensity") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
#USE
use.hist4 <- ggplot(data=peak2014spring, aes(x=use)) + geom_histogram(color="black", fill="blue", bins = 70) + xlab("Spring Use") + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
multiplot(gen.hist4, hum.hist4, appTEMP.hist4, cc.hist4, dewPOINT.hist4, temp.hist4, vis.hist4, press.hist4, windBearing.hist4, precipProb.hist4, windSpeed.hist4, pricipIntensity.hist4, cols=2)
```

MODEL
```{r}
spring1<-lm(gen~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb+use, data=peak2014spring)
summary(spring1) 
qqnorm(spring1$residuals)
anova(spring1) #icon, hum, press, windSpeed, cc, windBearing,pricipIntensity, precipProb, use... R^2= .2505

both4 = step(spring1, direction = "both", trace=FALSE ) 
summary(both4) #temp, iconclear-night,iconpartly-cloudy-night, iconrain, vis, sumDrizzle, sumMostly Cloudy, press, windSpeed, cc, windBearing, dewPoint, precipProb, use... R^2=.2648
backwards4 = step(spring1, direction = "backward", trace=FALSE ) 
summary(backwards4) #temp, iconclear-night,iconpartly-cloudy-night, iconrain, vis, sumDrizzle, sumMostly Cloudy, press, windSpeed, cc, windBearing, dewPoint, precipProb, use... R^2=.2648
forward4 = step(spring1, direction = "forward", trace=FALSE ) 
summary(forward4) #iconclear-night, iconrain, vis, sumMostly Cloudy, press, cc, windBearing, precipProb, use... R^2=.2614

spring_final<-lm(gen~temp+icon+ vis+sum+ press+windSpeed+ cc+ windBearing+ dewPoint+ precipProb+ use, data=peak2014spring)
pairs(gen~temp+ vis+ press+windSpeed+ cc+ windBearing+ dewPoint+ precipProb+ use, data=peak2014spring)
#vif(spring_final) #this code doesn't work again
alias(spring_final) #sumClear/iconcloudy, sumClear/iconfog, sumOvercast/iconcloudy,sumClear/iconpartly-cloudy-day,sumClear/iconpartly-cloudy-night, sumClear/iconrain, sumPartly Cloudy/iconpartly-cloudy-day, sumPartly Cloudy/iconpartly-cloudy-night, sumRain/iconrain, sumClear/iconwind, sumPartly Cloudy/sumMostly Cloudy, sumRain/sumDrizzle, sumRain/sumLight Rain seem to be highly correlated. 
qqnorm(spring_final$residuals)

#CHECKING ASSUMPTIONS
#the residuals vs fitted plot looks really bad...they clearly have some kind of trend
errors4<-ggplot(fortify(spring_final), aes(x=.fitted, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Fitted Values")+ylab("Residuals") +
      ggtitle("Residual vs Fitted Plot")
#check for non-constant variance
variance4<-ggplot(data=spring_final, aes(x=.resid)) + 
      geom_histogram(binwidth=0.5, fill="blue") +
      xlab("Residuals") +
      ggtitle("Distribution of Residuals")
```


################## COMPARING PLOTS ##################


#### DESCRIPTIVES

Generation
```{r}
multiplot(gen.hist, gen.hist2, gen.hist3, gen.hist4, cols=2)
```

Humidity
```{r}
multiplot(hum.hist, hum.hist2, hum.hist3, hum.hist4, cols=2)
```

appTEMP
```{r}
multiplot(appTEMP.hist, appTEMP.hist2, appTEMP.hist3, appTEMP.hist4, cols=2)
```

DewPoint
```{r}
multiplot(dewPOINT.hist, dewPOINT.hist2, dewPOINT.hist3, dewPOINT.hist4, cols=2)
```

TEMP
```{r}
multiplot(temp.hist, temp.hist2, temp.hist3, temp.hist4, cols=2)
```

Visibility
```{r}
multiplot(vis.hist,vis.hist2, vis.hist3, vis.hist4, cols=2)
```

Pressure
```{r}
multiplot(press.hist,press.hist2,press.hist3,press.hist4, cols=2)
```

Wind Bearing
```{r}
multiplot(windBearing.hist, windBearing.hist2, windBearing.hist3, windBearing.hist4, cols=2)
```

Precipitation Probability
```{r}
multiplot(precipProb.hist, precipProb.hist2, precipProb.hist3, precipProb.hist4, cols=2)
```

Wind Speed
```{r}
multiplot(windSpeed.hist, windSpeed.hist2, windSpeed.hist3, windSpeed.hist4, cols=2)
```

Precipitation Intensity
```{r}
multiplot(pricipIntensity.hist, precipProb.hist2, precipProb.hist3, precipProb.hist4, cols=2)
```

Use
```{r}
multiplot(use.hist, use.hist2, use.hist3, use.hist4, cols=2)
```

##### MODELS

```{r}
#want to take out use because it doesn't make sense. Also will probably take out sum or icon
#winter model has highest adj R^2
summary(w_final) #used anova, uses temp + icon + hum + vis + press + windBearing + use
summary(f_final) #used forward, uses temp + sum + cc + windBearing + dewPoint + precipProb + use
summary(summer_final) #used both/backwards, uses icon + hum + sum + appTemp + cc + windBearing +  pricipIntensity + dewPoint + precipProb + use
summary(spring_final) #used both/backwards, uses temp + icon + vis + sum + press + windSpeed + cc + windBearing + dewPoint + precipProb + use

#Normality of QQ Plot-they all look fine
qqnorm(w_final$residuals)
qqnorm(f_final$residuals)
qqnorm(summer_final$residuals)
qqnorm(spring_final$residuals)

#Correlation of errors
multiplot(errors1,errors2,errors3,errors4, cols=2) #winter,summer,fall,spring read across

#Constant Variance
multiplot(variance1,variance2,variance3,variance4, cols=2) #seems that all the season are distributed normally except winter
```

########## General Comments #############

-Seems that backwards and both selection produce the same adj R^2 for every season.

- in the winter, the specific variables are from the anova rather than the other 3 variable selection methods
-had to run an extra "alias" command for summer and spring. Maybe this is because there should be an interaction term? Seems that the variable sum and icon might say similar things. Tried taking out one of the variables and it didn't help. Also, tried the interaction term but didn't help. Have to look more into this.

-The winter model does not include sum but does include icon whereas every other season has both sum and icon

-summer model is only one that doesn't keep temperature in the model

################
THIS WAS MY HOUR LONG TANGENT TRYING TO GET THE ERRORS TO NOT BE CORRELATED...

INCLUDING TIME AND DATE (QUESTION: DID WE AVERAGE THE DATA FOR EACH HOUR??)

```{r}
peak2014t<- dplyr::select(homeC2014DAY,
                  temp:gen,
                  season,
                  time,
                  date
)
peak2014winter_t<-filter(peak2014t,season=="winter")
#looked at full model
w4<-lm(gen~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb+use+date+time, data=peak2014winter_t)
summary(w4) #R^2 is slightly better
anova(w4) #keeps temp, icon, hum, vis, press, windBearing, use, date, time
both = step(w4, direction = "both", trace=FALSE ) 
summary(both) #has an error 

#keeps temp, vis, press, cc, windBearing, dewPoint, use
backwards = step(w1, direction = "backward", trace=FALSE ) 
summary(backwards)
forward = step(w1, direction = "forward", trace=FALSE ) 
summary(forward) 
#check for non-linearity of data
res_plot<-ggplot(w2_2, aes(x=.fitted, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Fitted Values")+ylab("Residuals") +
      ggtitle("Residual vs Fitted Plot")
#correlation of error terms
error_correlation<-ggplot(w4, aes(x=time, y=.resid))+geom_point() +
      geom_smooth(se=FALSE)+geom_hline(yintercept=0, col="red", linetype="dashed") +
      xlab("Time")+ylab("Residuals") +
      ggtitle("Residual vs Time Plot")
```


WITH TRANSFORMS

```{r}
#this code doesn't work but idk why
w3<-lm(log(gen)~temp+icon+hum+vis+sum+appTemp+press+windSpeed+cc+windBearing+pricipIntensity+dewPoint+precipProb+use, data=peak2014winter)
```
